---
title: "Figures - Appendix A"
author: "Beauchamp Nathéo"
date: "2023-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(miscset)
library(RPostgreSQL)
library(xtable)
library(factoextra)
library(FactoMineR)
library(corrplot)
```

```{r include=FALSE}
lapply(list.files("R", recursive = TRUE), function(x) source(file.path("R", x)))
lapply(list.files("R_report", recursive = TRUE), function(x) source(file.path("R_report", x)))
```

```{r}
data_samsara <- sapply(c("calib_trees", "calib_plots"),
                         # "calib_soil", "calib_climate"),
                       Get_SamsaraDataset_from_PgAdmin,
                       Sys.getenv("LOGIN"),
                       Sys.getenv("PASSWORD"),
                       USE.NAMES = TRUE,
                       simplify = F)
```


```{r}
# Raw data
data_growth <- vroom("output/data_growth.csv") %>% 
  dplyr::filter(!species %in% c("Pinus canariensis", "Pinus radiata"))

data_mortality <- vroom("output/data_mortality.csv") %>% 
  dplyr::filter(!species %in% c("Pinus canariensis", "Pinus radiata"))
```


# Appendix

## Species shade tolerance

```{r}
data_shadetol_nandm <- vroom("data/data_Niinemets&Valladares_2006.csv", show_col_types = F)
data_shadetol_poorter <- vroom("data/data_Poorter_2012.csv", show_col_types = F)

species_shadetol <- c("Abies alba","Acer campestre","Acer pseudoplatanus","Alnus glutinosa","Betula pendula","Betula pubescens","Carpinus betulus","Castanea sativa","Fagus sylvatica","Fraxinus excelsior","Juniperus thurifera","Larix decidua","Picea abies","Pinus halepensis","Pinus nigra","Pinus pinaster","Pinus pinea","Pinus sylvestris","Pinus uncinata","Populus spp.","Prunus avium","Pseudotsuga menziesii","Quercus faginea","Quercus ilex","Quercus petraea","Quercus pubescens","Quercus pyrenaica","Quercus robur","Quercus suber","Robinia pseudoacacia","Salix spp.","Sorbus spp.","Tilia spp.")

# Bind both datasets
data_shadetol <- dplyr::bind_rows(
  data_shadetol_nandm %>% 
    dplyr::select(species = Species, shadetol = shade_tolerance.mean),
  data_shadetol_poorter %>% 
    dplyr::filter(species != "Pinus uncinata") # Not the same value as the dataset from Niinemets and Valladares
) %>% 
  dplyr::filter(species %in% unique(species_shadetol)) %>% 
  dplyr::distinct()

# Add lacking species
data_shadetol <- data_shadetol %>% 
  add_row(
    species = c("Populus spp.", "Juniperus thurifera", "Tilia spp.", "Salix spp.", "Sorbus spp."),
    shadetol = c(mean(data_shadetol_nandm$shade_tolerance.mean[grep("Populus", data_shadetol_nandm$Species)]),
                 mean(data_shadetol_nandm$shade_tolerance.mean[grep("Juniperus", data_shadetol_nandm$Species)]),
                 mean(data_shadetol_nandm$shade_tolerance.mean[grep("Tilia", data_shadetol_nandm$Species)]),
                 mean(data_shadetol_nandm$shade_tolerance.mean[grep("Salix", data_shadetol_nandm$Species)]),
                 mean(data_shadetol_nandm$shade_tolerance.mean[grep("Sorbus", data_shadetol_nandm$Species)]))
  ) %>% 
  dplyr::arrange(shadetol) %>% 
  dplyr::mutate(species = factor(species, levels = species))

ggplot(data_shadetol, aes(y = species, x = shadetol)) +
  geom_point()
```


# Overview

## Total number of individuals and plots

### Growth
```{r}
print(paste("Number of individuals:", nrow(data_growth)))
print(paste("Number of plots:", length(unique(data_growth$plotcode))))
```

### Mortality
```{r}
print(paste("Number of living individuals:", 
            nrow(data_mortality %>% dplyr::filter(!dead))))
print(paste("Number of dead and living individuals:", 
            nrow(data_mortality %>% dplyr::filter(dead == T))))
print(paste("Number of plots:", length(unique(data_mortality$plotcode))))
```


## Number of individuals per species

```{r, out.width="150%"}
# Number of individuals
data_occ <- data_growth %>% 
  dplyr::group_by(species) %>% 
  dplyr::summarise(occ_trees = n(),
                   occ_plots = length(unique(plotcode))) %>%
  dplyr::ungroup() %>% 
  dplyr::arrange(desc(occ_trees)) %>% 
  dplyr::mutate(species = factor(species, levels = species),
                lab_str = paste(formatC(occ_trees, format="d", big.mark=","),
                                formatC(occ_plots, format="d", big.mark=","),
                                sep = "\n")) 
  
ggplot(data_occ, aes(y = occ_trees, x = species)) +
  geom_histogram(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7)) +
  geom_text(aes(label = lab_str, y = occ_trees+10000), size = 1.35) +
  theme(panel.grid = element_blank(),
        axis.ticks = element_line(linewidth = 0.6),
        axis.line = element_line(linewidth = 0.4),
        axis.title.x = element_blank()) +
  ylab("number of individuals") +
  scale_y_continuous(labels = c("0", "100,000", "200,000", "300,000"))
```

## Species distribution across Europe

```{r, out.width = "150%", out.height = "150%"}
plot_list <- setNames(vector("list", length(unique(data_growth$species))),
                      sort(unique(data_growth$species)))
for (sp in unique(data_growth$species)) {
  data_tmp <- data_growth %>% 
    dplyr::filter(species == sp) %>% 
    dplyr::select(plotcode, longitude, latitude) %>% 
    dplyr::distinct() %>% 
    dplyr::mutate(id = row_number())
  
  plot_list[[sp]] <- Plot_EuropeMap(data_tmp, "id", color_grad = c("darkolivegreen", "darkolivegreen")) +
      theme(legend.position = "none", 
            plot.subtitle = element_text(hjust = 0.5, size = 8)) +
      labs(subtitle = sp)
}

cowplot::plot_grid(plotlist = plot_list)
```


# Climatic data

## Correction of temperatures with altitude

```{r}
data_climate_2018 <- vroom("data/data_climatic_rescaled_2018.csv", show_col_types = F)
```

```{r}
# Plot difference of altitude between Chelsa and observed plot
plot_diff_alt <- data_climate_2018 %>%
 dplyr::select(plotcode, altitude, alt_wc) %>%
 tidyr::drop_na() %>%
   dplyr::mutate(diff_altitude = abs(altitude - alt_wc)) %>%

   ggplot(aes(x = diff_altitude)) +
   geom_histogram() +
   xlim(c(0,500)) +
   ylab("number of plots") +
   xlab("absolute difference between altitudes (in m)") +
   theme_minimal() +
   theme(axis.ticks = element_line(linewidth = 0.6),
         axis.line = element_line(linewidth = 0.6),
         panel.grid = element_blank(),
         axis.title = element_text(size = 7))


 # Plot lapserates
 data_lr <- data_climate_2018 %>%
   dplyr::select(plotcode, contains("lr_")) %>%
   tidyr::pivot_longer(!plotcode,
                       names_pattern = "lr_(.*)_2018",
                       names_to = "month",
                       values_to = "lr") %>%
   dplyr::mutate(month = factor(month, levels = c(paste0("0", 1:9), 10:12)),
                 # Do not forget to rescale lapserate (still in Chelsa DB format)
                 lr = 0.1 * lr,
                 # Convert into lapserate per 100m
                 lr_per_100m = lr * 100
                 )

 print("monthly mean predicted local lapserate for the year 2018 is:")
 print(data_lr %>%
     dplyr::summarise(
       conf.low = attr(confint(lr_per_100m), "range")[[1]],
       qt025 = quantile(lr_per_100m, 0.025),
       mean = mean(lr_per_100m),
       qt975 = quantile(lr_per_100m, 0.975),
       conf.up = attr(confint(lr_per_100m), "range")[[2]]
     ))

 plot_lr <- data_lr %>%
   ggplot(aes(y = lr_per_100m, x = month)) +
   geom_hline(yintercept = -0.55, color = "salmon", linetype = "dashed", linewidth = 1.05) +
   geom_hline(yintercept = -0.65, color = "salmon", linetype = "dashed", linewidth = 1.05) +
   geom_boxplot(outlier.alpha = 0) +
   theme_minimal() +
   theme(axis.ticks = element_line(linewidth = 0.6),
         axis.line = element_line(linewidth = 0.6),
         panel.grid = element_blank(),
         axis.title = element_text(size = 7)) +
   ylim(c(-1, 0.2)) +
   ylab("lapserate (in °C per 100m)")


 # Plot temperature correction
 plot_tascorrect <- data_climate_2018 %>%
   dplyr::select(plotcode, altitude, contains("tas")) %>%
   tidyr::drop_na() %>%
   tidyr::pivot_longer(!c(plotcode, altitude),
                       names_pattern = "(.*)_(.*)_2018",
                       names_to = c("var", "month"),
                       values_to = "temperature") %>%
   dplyr::filter(var %in% c("tas", "tascorrect")) %>%
   tidyr::pivot_wider(names_from = "var",
                      values_from = "temperature") %>%
   dplyr::mutate(correction = tascorrect - tas,
                 altitude_class = case_when(
                   altitude < 500 ~ "0-500",
                   TRUE ~ "sup500"),
                 altitude_class = factor(altitude_class, levels = c("0-500", "sup500"))) %>%
   tidyr::drop_na() %>%

   ggplot(aes(y = abs(correction), x = month)) +
   geom_boxplot(outlier.alpha = 0) +
   ylim(c(0, 1.2)) +
   theme_minimal() +
   theme(axis.ticks = element_line(linewidth = 0.6),
         axis.line = element_line(linewidth = 0.6),
         panel.grid = element_blank(),
         axis.title = element_text(size = 7)) +
   facet_wrap(~altitude_class) +
   ylab("absolute temperature correction (in °C)")


 # Combine plots
 cowplot::plot_grid(

   cowplot::plot_grid(
     plot_diff_alt,
     plot_lr,
     nrow = 1,
     labels = c("A", "B"),
   label_size = 12),

   plot_tascorrect,

   ncol = 1,
   labels = c("", "C"),
   label_size = 12
 )
```

### Snow model

```{r}
data_snow <- data_samsara$calib_climate %>%
  dplyr::select(plotcode, starts_with(c("snostor_", "pr", "lpr_", "spr", "lw_"))) %>%
  dplyr::select(plotcode, contains("chelsaperiod"),
                -contains(c(".gs", "_DD"))) %>%
  tidyr::drop_na() %>%
  tidyr::pivot_longer(
    !plotcode,
    names_pattern = "(.*)\\.(.*)_chelsaperiod",
    names_to = c("var", "season"),
    values_to = "value"
  ) %>%
  tidyr::drop_na() %>%
  dplyr::left_join(data_samsara$calib_plots, by = "plotcode") %>% 
  dplyr::right_join(data_growth %>% dplyr::select(plotcode) %>% dplyr::distinct(), by = "plotcode")
```

```{r}
plot_rain_snow <- ggplot(data_snow %>%
         dplyr::filter(var %in% c("lpr_MM", "spr_MM")) %>%
         dplyr::mutate(precipitations = ifelse(var == "lpr_MM",
                                               "rain", "snow"),
                       value = value / 4,
                       season = factor(season, levels = c("winter", "spring",
                                                          "summer", "autumn"))),
       aes(y = value, x = country, color = precipitations)) +
  facet_wrap(~season, nrow = 1) +
  scale_color_manual(values = c("#3399ff", "#99ccff")) +
  geom_boxplot(outlier.alpha = 0) +
  ylab("precipitations (in mm/month)") +
  guides(color = guide_legend(title = "")) +
  ylim(0, 200) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(size = 5),
        axis.title.y = element_text(size = 7),
        axis.ticks = element_line(linewidth = 0.6),
        axis.line = element_line(linewidth = 0.6),
        panel.grid = element_blank(),
        axis.title = element_text(size = 7),
        strip.text.x = element_text(face = "bold"))

plot_diff_prec <- data_snow %>%
  dplyr::filter(var %in% c("pr", "lw_MM")) %>%
  dplyr::mutate(value = value / 4) %>%
  tidyr::pivot_wider(names_from = "var",
                     values_from = "value") %>%
  dplyr::mutate(diff = pr - lw_MM,
                season = factor(season, levels = c("winter", "spring",
                                                   "summer", "autumn"))) %>%
  ggplot(aes(y = diff, x = country)) +
  facet_wrap(~season, nrow = 1) +
  geom_hline(yintercept = 0, color = "salmon", linetype = "dashed", linewidth = 1.05) +
  geom_boxplot(outlier.alpha = 0) +
  ylim(c(-40, 40)) +
  ylab("pr - lw (in mm/month)") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5),
        axis.title.y = element_text(size = 7),
        axis.ticks = element_line(linewidth = 0.6),
        axis.line = element_line(linewidth = 0.6),
        panel.grid = element_blank(),
        axis.title = element_text(size = 7),
        strip.text.x = element_text(face = "bold"))

cowplot::plot_grid(
  plot_rain_snow,
  plot_diff_prec,
  ncol = 1,
  labels = c("A", "B")
)
```


### Soil water holding capacity

```{r}
data_soil_plots <- data_samsara$calib_soil %>%
  dplyr::left_join(data_samsara$calib_plots, by = "plotcode") %>% 
  dplyr::right_join(data_growth %>% dplyr::select(plotcode) %>% dplyr::distinct(), by = "plotcode")

plot_silt <- data_soil_plots %>%
  Plot_EuropeMap("silt",
                 color_grad = c("white", "black")) +
  labs(subtitle = "silt") +
  theme(plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top") +
  guides(color = guide_colorbar(title="composition (%)",
                                title.theme = element_text(size = 11),
                                title.position = "top",
                                label.theme = element_text(size = 8)))

legend_perc <- get_legend(plot_silt)
plot_silt <- plot_silt + theme(legend.position = "none")

plot_sand <- data_soil_plots %>%
  Plot_EuropeMap("sand",
                 color_grad = c("white", "black")) +
  labs(subtitle = "sand") +
  theme(plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")


plot_clay <- data_soil_plots %>%
  Plot_EuropeMap("clay",
                 color_grad = c("white", "black")) +
  labs(subtitle = "clay") +
  theme(plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")

plot_soc <- data_soil_plots %>%
  Plot_EuropeMap("soc",
                 color_grad = c("white", "black")) +
  theme(legend.position = "top") +
  guides(color = guide_colorbar(title="organic carbon (g/kg)",
                                title.theme = element_text(size = 11),
                                title.position = "top",
                                label.theme = element_text(size = 8)))

plot_rootingdepth <- data_soil_plots %>%
  Plot_EuropeMap("rooting_depth",
                 color_grad = c("#e7d89d", "#663300")) +
  theme(legend.position = "top") +
  guides(color = guide_colorbar(title="rooting depth (m)",
                                title.theme = element_text(size = 11),
                                title.position = "top",
                                label.theme = element_text(size = 8)))

plot_swhc <- Compute_SWHCTot(
  data_soil_plots,
  c("0-5cm", "5-15cm", "15-30cm", "30-60cm", "60-100cm", "100-200cm")
) %>%
  dplyr::left_join(data_samsara$calib_plots, by = "plotcode") %>%
  Plot_EuropeMap("swhcTot", color_grad = c("#e7d89d", "#0a01fc")) +
  theme(legend.position = "top") +
  guides(color = guide_colorbar(title="SWHC (mm)",
                                title.theme = element_text(size = 12),
                                title.position = "top",
                                label.theme = element_text(size = 10)))

cowplot::plot_grid(
  cowplot::plot_grid(
    cowplot::plot_grid(
      legend_perc,
      cowplot::plot_grid(
        plot_silt, plot_sand, plot_clay,
        nrow = 1, labels = c("1a", "1b", "1c")),
      ncol = 1, rel_heights = c(1, 3)
    ),
    cowplot::plot_grid(
      plot_soc, plot_rootingdepth,
      nrow = 1, labels = c("2", "3")
    ),
    ncol = 1
  ),
  plot_swhc,
  nrow = 1, labels = c("", "4")
)
```

# M&M figure - Climate

```{r}
plot_sgdd <- data_growth %>%
  Plot_EuropeMap("sgdd", c("#82CEF9", "#F1895C")) +
  labs(title = "temperature") +
  theme(# legend.position = "right",
        # plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
        # legend.title = element_text(size = 8),
        # legend.text = element_text(size = 7),
        legend.text = element_blank(),
        plot.title = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom")

plot_aet2pet <- data_growth %>%
  Plot_EuropeMap("aet2pet", c("#b01111", "#EEE2DF")) +
  labs(title = "water availability") +
  theme(
        # legend.position = "right",
        # plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
        # legend.title = element_text(size = 8),
        # legend.text = element_text(size = 7),
        legend.text = element_blank(),
        plot.title = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom"
        )

cowplot::plot_grid(
  plot_sgdd,
  plot_aet2pet,
  nrow = 1
)
```



# Light competition index variability

```{r}
data_light <- Get_SamsaraDataset_from_PgAdmin("calib_compet",
                      Sys.getenv("LOGIN"),
                      Sys.getenv("PASSWORD"))
```

```{r}
ggplot(data_light %>% drop_na(), aes(y = lci_sd, x = lci_mean)) +
 geom_point(alpha = 0.1) +
 geom_smooth(color = "salmon") +
 xlim(c(0, 1)) +
 ylim(c(0, 1)) +
 ylab("Standard deviation of estimated lci of equivalent trees") +
 xlab("Mean lci of equivalent trees") +
 theme_minimal()
```

```{r}
data_growth %>%
  dplyr::left_join(data_light, by = "treecode") %>% 
  dplyr::filter(species %in% c("Pinus sylvestris", "Quercus petraea", "Fagus sylvatica")) %>%
  dplyr::mutate(group_light = as.factor((lci_mean*100) %/% 5)) %>%
  dplyr::group_by(species, group_light) %>%
  dplyr::summarise(lci_mean = mean(lci_mean),
                   lciSD_mean = mean(lci_sd),
                   upper = lci_mean + lciSD_mean,
                   lower = lci_mean - lciSD_mean) %>% 

  ggplot(aes(y = lci_mean, x = lci_mean)) +
  facet_wrap(~species) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_point(color = "salmon") +
  geom_pointrange(aes(ymin = lower, ymax = upper), color = "black") +
  theme_minimal() +
  theme(axis.ticks = element_line(linewidth = 0.6),
        axis.line = element_line(linewidth = 0.6),
        panel.grid = element_blank()) +
  xlab("Mean lci for each class") +
  ylab("Mean lci for each class")
```



# Links between variables

## Global correlations by individuals

```{r}
M_cor <- cor(data_growth %>%
               dplyr::select(dbh, sgdd, aet2pet, lci, bat, bal))

# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
p.mat <- cor.mtest(data_growth %>%
               dplyr::select(dbh, sgdd, aet2pet, lci, bat, bal))

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M_cor, method="color", col=col(200),
         type="upper", order="hclust",
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.01, insig = "blank",
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
         )
```


## Correlation between climate variables

### All species

```{r}
data_pca <- data_growth %>%
  dplyr::select(plotcode, country, sgdd, aet2pet) %>%
  dplyr::distinct(plotcode, .keep_all = T)

cor(data_pca$sgdd, data_pca$aet2pet)
data_pca %>%
  dplyr::group_by(country) %>%
  dplyr::summarise(cor = cor(aet2pet, sgdd))

res.pca <- PCA(data_pca[,c("sgdd", "aet2pet")], scale.unit = T, ncp = 2, graph = FALSE)

fviz_pca_var(res.pca, col.var = "black")

ind <- get_pca_ind(res.pca)
data_pca$climate <- ind$coord[,1]
data_pca$cos2 <- ind$cos2[,1]


data_pca %>%
ggplot(aes(y = cos2, x = country)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.ticks = element_line(linewidth = 0.6),
        axis.line = element_line(linewidth = 0.6),
        panel.grid = element_blank())

data_growth %>%
  dplyr::select(plotcode, country, aet2pet, sgdd) %>%
  dplyr::distinct() %>%

  ggplot(aes(y = aet2pet, x = sgdd)) +
  geom_hex() +
  geom_smooth()

data_growth %>%
  dplyr::select(plotcode, country, aet2pet, sgdd) %>%
  dplyr::distinct() %>%

  ggplot(aes(y = aet2pet, x = sgdd)) +
  # geom_point(alpha = 0.1) +
  facet_wrap(~country) +
  geom_smooth()
```


```{r, out.width="150%", out.height="150%"}
data_climate_shadetol <- data_growth %>%
         dplyr::select(species, aet2pet, sgdd) %>%
         dplyr::group_by(species) %>%
         dplyr::summarise_all(list("mean" = mean,
                                   "qt1" = ~quantile(., 0.1),
                                   "qt9" = ~quantile(., 0.9))) %>%
         dplyr::right_join(data_shadetol, by = "species") %>%
         dplyr::arrange(species) %>%
         dplyr::mutate(id = row_number())

plot_climate_shadetol <- ggplot(data_climate_shadetol,
                                aes(y = aet2pet_mean, x = sgdd_mean, color = shadetol)) +
  geom_point() +
  scale_color_gradient2(midpoint = mean(data_shadetol$shadetol), space = "Lab",
                        low = "goldenrod1", mid = "orange3", high = "lightsteelblue4") +
  geom_pointrange(mapping = aes(ymin = aet2pet_qt1, ymax = aet2pet_qt9)) +
  geom_pointrange(mapping = aes(xmin = sgdd_qt1, xmax = sgdd_qt9)) +
  xlab("Sum of growing degree days sgdd (°C)") +
  ylab("Ratio of actual to potential evapotranspiration aet2pet") +
  labs(color = "shadetol") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.title = element_text(size = 9),
        legend.position = "right") +
  geom_text(mapping = aes(label = id), color = "black", size = 1.5)

species_climate_shadetol <- paste(paste(data_climate_shadetol$id,
                                  data_climate_shadetol$species,
                                  sep = " : "), collapse = "\n")

cowplot::plot_grid(
   ggdraw() + draw_label(species_climate_shadetol, size = 8, hjust=0, x=0.05),
   plot_climate_shadetol,
   nrow = 1, rel_widths = c(1,4)
)
```


### Within-species

```{r}
data_climate_within <- data_growth %>%
  dplyr::group_by(species) %>%
  dplyr::summarise(cor = cor(sgdd, aet2pet)) %>%
  dplyr::arrange(cor) %>%
  tidyr::pivot_wider(names_from = "species", values_from = "cor")

xtable(data_climate_within)
```


```{r}
data_growth %>%
  dplyr::group_by(species) %>%
  dplyr::summarise(cor = cor(sgdd, aet2pet)) %>%
  dplyr::summarise(cor_mean = mean(abs(cor)))
```


```{r}
data_growth %>%
  dplyr::select(species, aet2pet, sgdd) %>%

  ggplot(aes(y = aet2pet, x = sgdd)) +
  facet_wrap(~species) +
  geom_smooth(method = "lm")
```


## Correlation between competition variables

```{r}
r2_lci_compet <- dplyr::left_join(

  sapply(unique(data_growth$species), function(x) {
    round(summary(
      lm(lci ~ bal, data_growth %>%  dplyr::filter(species == x)
      ))$r.squared, 2)
  }) %>%
    data.frame(species = names(.), bal = .),

  sapply(unique(data_growth$species), function(x) {
    round(summary(
      lm(lci ~ bat, data_growth %>%  dplyr::filter(species == x)
      ))$r.squared, 2)
  }) %>%
    data.frame(species = names(.), bat = .),

  by = "species") %>%

  dplyr::left_join(

    sapply(unique(data_growth$species), function(x) {
      round(summary(
        lm(lci ~ bat*dbh, data_growth %>%  dplyr::filter(species == x)
        ))$r.squared, 2)
    }) %>%
      data.frame(species = names(.), batXdbh = .),

    by = "species") %>%

  dplyr::arrange(batXdbh)

print(paste("Mean R2 over species with bal is", round(mean(r2_lci_compet$bal), 2),
            "ranging from", round(min(r2_lci_compet$bal), 2),
            "(", r2_lci_compet$species[which.min(r2_lci_compet$bal)] ,")",
            "to", round(max(r2_lci_compet$bal), 2),
            "(", r2_lci_compet$species[which.max(r2_lci_compet$bal)] ,")",
            " - mean R2 over species with bat is", round(mean(r2_lci_compet$bat), 2),
            "ranging from", round(min(r2_lci_compet$bat), 2),
            "(", r2_lci_compet$species[which.min(r2_lci_compet$bat)] ,")",
            "to", round(max(r2_lci_compet$bat), 2),
            "(", r2_lci_compet$species[which.max(r2_lci_compet$bat)] ,")",
            " - mean R2 over species with batXdbh is", round(mean(r2_lci_compet$batXdbh), 2),
            "ranging from", round(min(r2_lci_compet$batXdbh), 2),
            "(", r2_lci_compet$species[which.min(r2_lci_compet$batXdbh)] ,")",
            "to", round(max(r2_lci_compet$batXdbh), 2),
            "(", r2_lci_compet$species[which.max(r2_lci_compet$batXdbh)] ,")"))

r2_lci_compet_mat <- matrix(
  c(r2_lci_compet$batXdbh,
    r2_lci_compet$bat,
    r2_lci_compet$bal),
  nrow = 3, byrow = TRUE,
  dimnames = list(c("batXdbh", "bat", "bal"), r2_lci_compet$species)
)

xtable(r2_lci_compet_mat)
```



## Correlation between light competition index and climate


### Correlation by species

```{r}
cor_lci_climate <- data_growth %>%
  dplyr::group_by(species) %>%
  dplyr::summarise(cor_sgdd = cor(lci, sgdd),
                   cor_aet2pet = cor(lci, aet2pet)) %>%
  dplyr::arrange(cor_aet2pet)

print(paste("Mean Pearson index over species between lci and sgdd is", round(mean(abs(cor_lci_climate$cor_sgdd)), 2),
            "ranging from", round(min(cor_lci_climate$cor_sgdd), 2),
            "(", cor_lci_climate$species[which.min(cor_lci_climate$cor_sgdd)] ,")",
            "to", round(max(cor_lci_climate$cor_sgdd), 2),
            "(", cor_lci_climate$species[which.max(cor_lci_climate$cor_sgdd)] ,")",
            "- mean Pearson index over species between lci and aet2pet is", round(mean(abs(cor_lci_climate$cor_aet2pet)), 2),
            "ranging from", round(min(cor_lci_climate$cor_aet2pet), 2),
            "(", cor_lci_climate$species[which.min(cor_lci_climate$cor_aet2pet)] ,")",
            "to", round(max(cor_lci_climate$cor_aet2pet), 2),
            "(", cor_lci_climate$species[which.max(cor_lci_climate$cor_aet2pet)] ,")"))

cor_lci_climate_mat <- matrix(
  c(cor_lci_climate$cor_aet2pet,
    cor_lci_climate$cor_sgdd),
  nrow = 2, byrow = TRUE,
  dimnames = list(c("aet2pet", "sgdd"), cor_lci_climate$species)
)

xtable(cor_lci_climate_mat)
```


### Quantiles regressions

```{r, out.width = "150%", out.height = "150%"}
data_growth %>%
  ggplot(aes(y = lci, x = aet2pet)) +
  facet_wrap(~species) +
  geom_quantile(quantiles = c(0.1, 0.5, 0.9), formula = y~x) +
  geom_hline(yintercept = 0.1, color = "salmon") +
  geom_hline(yintercept = 0.9, color = "salmon") +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 6),
        axis.text = element_text(size = 6),
        axis.text.x = element_text(angle = 45, hjust = 0.5))

data_growth %>%
  ggplot(aes(y = lci, x = sgdd)) +
  facet_wrap(~species) +
  geom_quantile(quantiles = c(0.1, 0.5, 0.9), formula = y~x) +
  geom_hline(yintercept = 0.1, color = "salmon") +
  geom_hline(yintercept = 0.9, color = "salmon") +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 6),
        axis.text = element_text(size = 6),
        axis.text.x = element_text(angle = 45, hjust = 0.5))
```